 

### 2019-03-03 Manuscript/SupplementaryMaterial.Rmd
Manuscript Update

Both cylindrical models introduced here are estimated using Markov Chain Monte
Carlo (MCMC) methods based on @nunez2011bayesian, @wang2012directional and
@hernandez2016general for the regression of the circular outcome. A detailed
description of the Bayesian estimation and MCMC samplers can be found in the
Supplementary Material.

@@

This model is an extension of the cylindrical model introduced in
@abe2017tractable to the regression context. The joint density of \(\Theta\) and
\(Y\), in this model defined only on the positive real half-line \([0, +
\infty)\), reads
\begin{equation}\label{WeiSSVMdensity}
f(\theta, y) = \frac{\alpha\nu^\alpha}{2\pi\cosh(\kappa)}
                 (1 +\lambda\sin(\theta - \mu))
                 y^{\alpha-1}
                 \exp[-(\nu y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu))],
\end{equation}
\noindent where \(\alpha > 0\) is a linear shape parameter, \(\kappa > 0\) and
\(\lambda \in [-1, 1]\) are circular concentration and skewness parameters with
\(\kappa\) also regulating the circular-linear dependence. Our modification
occurs at the level of the linear scale parameter \(\nu>0\) and circular
location parameter \(\mu\in [0, 2\pi)\), both of which we express in terms of
covariates: \(\nu_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
\(\mu_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\). The
parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression coefficients
\(\gamma_j \in (-\infty, +\infty)\) for the prediction of \(y\) where \(j = 0,
\dots, q\) and \(\nu_0\) is the intercept. The parameter \(\beta_0 \in [0,
2\pi)\) is the intercept and \(\boldsymbol{\beta}\) is a vector of \(p\)
regression coefficients \(\beta_j \in (-\infty, +\infty)\) for the prediction of
\(\theta\) where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a
vector of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
is a vector of predictor values for the prediction of \(\theta\). In a
regression setup the outcome vector \((\theta_i, y_i)^t\) for each individual is
generated independently from the modified
density \eqref{WeiSSVMdensity}.\newline
\indent As in @abe2017tractable, the conditional distribution of \(Y\) given
\(\Theta=\theta\) is a Weibull distribution with shape \(\alpha\) and scale
\(\nu(1-\tanh(\kappa)\cos(\theta - \mu))^{1/\alpha}\) and the conditional
distribution of \(\Theta\) given \(Y=y\) is a sine skewed von Mises distribution
with location parameter \(\mu\) and concentration parameter \((\nu
y)^\alpha\tanh(\kappa)\). The log-likelihood for this model equals
\begin{align}\label{WeiSSVMLikelihood}
l(\alpha, \boldsymbol{\gamma}, \lambda, \kappa, \boldsymbol{\beta}) 
   &= n[\ln(\alpha) - \ln(2\pi\cosh(\kappa))] + \alpha \sum^{n}_{i = 1} \boldsymbol{x}_i^t\boldsymbol{\gamma} \nonumber\\
   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - \beta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta}))) 
   +(\alpha-1)\sum^{n}_{i = 1} \ln(y_i) \nonumber\\
   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - \beta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta}))).\nonumber
\end{align}
\noindent We can use numerical optimization (Nelder-Mead) to find solutions for
the maximum likelihood (ML) estimates for the parameters of the model.

@@

\indent We estimate the model using MCMC methods. A detailed description of these
methods is given in the Supplementary Material.

@@

\indent We use the following loglikelihoods for the
computation of the PLSL in the teacher data:

@@

$$l(y) = \log(1) - \log(\sqrt{2\pi\sigma^2}) + \sum(\hat{y}_i-(\gamma_0 + \gamma_{cos}\cos(\theta_i)r_i +
\gamma_{sin}\sin(\theta_i)r_i + \gamma_1\text{SE}_i))^2/2\sigma^2$$

$$l(\theta) = \log(1) - \log(2\pi) + \sum -0.5\hat{\mu}_i^2 - 0.5(r_i^2 - 2r_iu_i^t\hat{\mu}_i)$$
where $u_i = (cos\theta_i, \sin \theta_i)$ and $\hat{\mu}_i = (\beta_0^{I} + \beta_0^{I}\text{SE}_i, \beta_0^{II} + \beta_0^{II}\text{SE}_i)^t$.

@@

$$l(y) = \log(1) - \log(\sqrt{2\pi\sigma^2}) + \sum(\hat{y}_i-(\gamma_0 + \gamma_{cos}\cos(\theta_i)r_i +
\gamma_{sin}\sin(\theta_i)r_i + \gamma_1\text{SE}_i))^2/2\sigma^2$$

@@

$$l(y) = \log\alpha + \sum\log h_i^\alpha + \sum\log y_i^{\alpha - 1} - \sum(h_iy_i)^\alpha$$

where $h_i = \exp\{\hat{y}_i\}\{1-\tanh(\kappa)\cos(\theta_i - \hat{\theta}_i)\}^{1/\alpha}$, $\hat{y}_i = \gamma_0 + \gamma_1\text{SE}_i$ and $\hat{\theta}_i = \beta_0 +  2\tan^{-1}(\beta_1\text{SE}_i))$ .

$$l(\theta) = \log(1) - \sum\log 2\pi I_0(c_i) + \sum\log\{1 + \lambda \sin(\theta_i - \hat{\theta}_i\} + \sum c_i\cos(\theta_i - \hat{\theta}_i)$$

where $c_i = y_i^{\alpha}\exp(\hat{y}_i)^{\alpha}\tanh\kappa$, and $I_{0}$ is a Bessel function of order $0$.

@@

\end{enumerate}

\section*{References}
<div id="refs"></div>

>>>>

Both cylindrical models introduced here are estimated using Markov Chain Monte
Carlo (MCMC) methods based on @nunez2011bayesian, @wang2012directional and
@hernandez2016general for the regression of the circular outcome.

@@

This model is an extension of the cylindrical model introduced in
@abe2017tractable to the regression context. The joint density of \(\Theta\) and
\(Y\), in this model defined only on the positive real half-line \([0, +
\infty)\), reads
\begin{equation}\label{WeiSSVMdensity}
f(\theta, y) = \frac{\alpha\nu^\alpha}{2\pi\cosh(\kappa)}
                 (1 +\lambda\sin(\theta - \mu))
                 y^{\alpha-1}
                 \exp[-(\nu y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu))],
\end{equation}
\noindent where \(\alpha > 0\) is a linear shape parameter, \(\kappa > 0\) and
\(\lambda \in [-1, 1]\) are circular concentration and skewness parameters with
\(\kappa\) also regulating the circular-linear dependence. Our modification
occurs at the level of the linear scale parameter \(\nu>0\) and circular
location parameter \(\mu\in [0, 2\pi)\), both of which we express in terms of
covariates: \(\nu_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
\(\mu_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\). The
parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression coefficients
\(\gamma_j \in (-\infty, +\infty)\) for the prediction of \(y\) where \(j = 0,
\dots, q\) and \(\nu_0\) is the intercept. The parameter \(\beta_0 \in [0,
2\pi)\) is the intercept and \(\boldsymbol{\beta}\) is a vector of \(p\)
regression coefficients \(\beta_j \in (-\infty, +\infty)\) for the prediction of
\(\theta\) where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a
vector of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
is a vector of predictor values for the prediction of \(\theta\). In a
regression setup the outcome vector \((\theta_i, y_i)^t\) for each individual is
generated independently from the modified
density \eqref{WeiSSVMdensity}.\newline
\indent As in @abe2017tractable, the conditional distribution of \(Y\) given
\(\Theta=\theta\) is a Weibull distribution with shape \(\alpha\) and scale
\(\nu(1-\tanh(\kappa)\cos(\theta - \mu))^{1/\alpha}\) and the conditional
distribution of \(\Theta\) given \(Y=y\) is a sine skewed von Mises distribution
with location parameter \(\mu\) and concentration parameter \((\nu
y)^\alpha\tanh(\kappa)\). The log-likelihood for this model equals
\begin{align}\label{WeiSSVMLikelihood}
l(\alpha, \boldsymbol{\gamma}, \lambda, \kappa, \boldsymbol{\beta}) 
   &= n[\ln(\alpha) - \ln(2\pi\cosh(\kappa))] + \alpha \sum^{n}_{i = 1} \boldsymbol{x}_i^t\boldsymbol{\gamma} \nonumber\\
   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - (\beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})))) 
   +(\alpha-1)\sum^{n}_{i = 1} \ln(y_i) \nonumber\\
   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - (\beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})))).\nonumber
\end{align}
\noindent We can use numerical optimization (Nelder-Mead) to find solutions for
the maximum likelihood (ML) estimates for the parameters of the model.

@@

@@

\indent We use the following (conditional) loglikelihoods for the
computation of the PLSL in the teacher data:

@@

$$l(y \mid \theta, r) = \log(1) - \log(\sqrt{2\pi\sigma^2}) + \sum(\hat{y}_i-(\gamma_0 + \gamma_{cos}\cos(\theta_i)r_i +
\gamma_{sin}\sin(\theta_i)r_i + \gamma_1\text{SE}_i))^2/2\sigma^2$$

$$l(\theta, r) = \log(1) - \log(2\pi) + \sum -0.5\hat{\mu}_i^2 - 0.5(r_i^2 - 2r_iu_i^t\hat{\mu}_i)$$
where $u_i = (cos\theta_i, \sin \theta_i)$ and $\hat{\mu}_i = (\beta_0^{I} + \beta_0^{I}\text{SE}_i, \beta_0^{II} + \beta_0^{II}\text{SE}_i)^t$.

@@

$$l(y \mid \theta, r) = \log(1) - \log(\sqrt{2\pi\sigma^2}) + \sum(\hat{y}_i-(\gamma_0 + \gamma_{cos}\cos(\theta_i)r_i +
\gamma_{sin}\sin(\theta_i)r_i + \gamma_1\text{SE}_i))^2/2\sigma^2$$

@@

$$l(y \mid \theta) = \log\alpha + \sum\log h_i^\alpha + \sum\log y_i^{\alpha - 1} - \sum(h_iy_i)^\alpha$$

where $h_i = \exp(\hat{y}_i)\{1-\tanh(\kappa)\cos(\theta_i - \hat{\theta}_i)\}^{1/\alpha}$, $\hat{y}_i = \gamma_0 + \gamma_1\text{SE}_i$ and $\hat{\theta}_i = \beta_0 +  2\tan^{-1}(\beta_1\text{SE}_i))$ .

$$l(\theta \mid y) = \log(1) - \sum\log 2\pi I_0(c_i) + \sum\log\{1 + \lambda \sin(\theta_i - \hat{\theta}_i\} + \sum c_i\cos(\theta_i - \hat{\theta}_i)$$

where $c_i = y_i^{\alpha}\exp(\hat{y}_i)^{\alpha}\tanh\kappa$, and $I_{0}$ is a modified Bessel function of order $0$.

@@

\end{enumerate}
\newpage
\section*{References}
<div id="refs"></div>

### 2019-03-03 Manuscript/SupplementaryMaterial.tex
Manuscript Update

Both cylindrical models introduced here are estimated using Markov Chain
Monte Carlo (MCMC) methods based on Nuñez-Antonio et al. (2011), Wang \&
Gelfand (2013) and Hernandez-Stumpfhauser et al. (2016) for the
regression of the circular outcome. A detailed description of the
Bayesian estimation and MCMC samplers can be found in the Supplementary
Material.

@@

This model is an extension of the cylindrical model introduced in Abe \&
Ley (2017) to the regression context. The joint density of \(\Theta\)
and \(Y\), in this model defined only on the positive real half-line
\([0, + \infty)\), reads \begin{equation}\label{WeiSSVMdensity}
f(\theta, y) = \frac{\alpha\nu^\alpha}{2\pi\cosh(\kappa)}
                 (1 +\lambda\sin(\theta - \mu))
                 y^{\alpha-1}
                 \exp[-(\nu y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu))],
\end{equation} \noindent where \(\alpha > 0\) is a linear shape
parameter, \(\kappa > 0\) and \(\lambda \in [-1, 1]\) are circular
concentration and skewness parameters with \(\kappa\) also regulating
the circular-linear dependence. Our modification occurs at the level of
the linear scale parameter \(\nu>0\) and circular location parameter
\(\mu\in [0, 2\pi)\), both of which we express in terms of covariates:
\(\nu_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
\(\mu_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\).
The parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression
coefficients \(\gamma_j \in (-\infty, +\infty)\) for the prediction of
\(y\) where \(j = 0, \dots, q\) and \(\nu_0\) is the intercept. The
parameter \(\beta_0 \in [0, 2\pi)\) is the intercept and
\(\boldsymbol{\beta}\) is a vector of \(p\) regression coefficients
\(\beta_j \in (-\infty, +\infty)\) for the prediction of \(\theta\)
where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a vector
of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
is a vector of predictor values for the prediction of \(\theta\). In a
regression setup the outcome vector \((\theta_i, y_i)^t\) for each
individual is generated independently from the modified density
\eqref{WeiSSVMdensity}.\newline \indent As in Abe \& Ley (2017), the
conditional distribution of \(Y\) given \(\Theta=\theta\) is a Weibull
distribution with shape \(\alpha\) and scale
\(\nu(1-\tanh(\kappa)\cos(\theta - \mu))^{1/\alpha}\) and the
conditional distribution of \(\Theta\) given \(Y=y\) is a sine skewed
von Mises distribution with location parameter \(\mu\) and concentration
parameter \((\nu y)^\alpha\tanh(\kappa)\). The log-likelihood for this
model equals \begin{align}\label{WeiSSVMLikelihood}
l(\alpha, \boldsymbol{\gamma}, \lambda, \kappa, \boldsymbol{\beta}) 
   &= n[\ln(\alpha) - \ln(2\pi\cosh(\kappa))] + \alpha \sum^{n}_{i = 1} \boldsymbol{x}_i^t\boldsymbol{\gamma} \nonumber\\
   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - \beta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta}))) 
   +(\alpha-1)\sum^{n}_{i = 1} \ln(y_i) \nonumber\\
   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - \beta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta}))).\nonumber
\end{align} \noindent We can use numerical optimization (Nelder-Mead) to
find solutions for the maximum likelihood (ML) estimates for the
parameters of the model.

@@

intercept and the \(g\) covariate values. \newline \indent We estimate
the model using MCMC methods. A detailed description of these methods is
given in the Supplementary Material.

@@

\end{enumerate}

\section*{References}

>>>>

Both cylindrical models introduced here are estimated using Markov Chain
Monte Carlo (MCMC) methods based on Nuñez-Antonio et al. (2011), Wang \&
Gelfand (2013) and Hernandez-Stumpfhauser et al. (2016) for the
regression of the circular outcome.

@@

This model is an extension of the cylindrical model introduced in Abe \&
Ley (2017) to the regression context. The joint density of \(\Theta\)
and \(Y\), in this model defined only on the positive real half-line
\([0, + \infty)\), reads \begin{equation}\label{WeiSSVMdensity}
f(\theta, y) = \frac{\alpha\nu^\alpha}{2\pi\cosh(\kappa)}
                 (1 +\lambda\sin(\theta - \mu))
                 y^{\alpha-1}
                 \exp[-(\nu y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu))],
\end{equation} \noindent where \(\alpha > 0\) is a linear shape
parameter, \(\kappa > 0\) and \(\lambda \in [-1, 1]\) are circular
concentration and skewness parameters with \(\kappa\) also regulating
the circular-linear dependence. Our modification occurs at the level of
the linear scale parameter \(\nu>0\) and circular location parameter
\(\mu\in [0, 2\pi)\), both of which we express in terms of covariates:
\(\nu_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
\(\mu_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\).
The parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression
coefficients \(\gamma_j \in (-\infty, +\infty)\) for the prediction of
\(y\) where \(j = 0, \dots, q\) and \(\nu_0\) is the intercept. The
parameter \(\beta_0 \in [0, 2\pi)\) is the intercept and
\(\boldsymbol{\beta}\) is a vector of \(p\) regression coefficients
\(\beta_j \in (-\infty, +\infty)\) for the prediction of \(\theta\)
where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a vector
of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
is a vector of predictor values for the prediction of \(\theta\). In a
regression setup the outcome vector \((\theta_i, y_i)^t\) for each
individual is generated independently from the modified density
\eqref{WeiSSVMdensity}.\newline \indent As in Abe \& Ley (2017), the
conditional distribution of \(Y\) given \(\Theta=\theta\) is a Weibull
distribution with shape \(\alpha\) and scale
\(\nu(1-\tanh(\kappa)\cos(\theta - \mu))^{1/\alpha}\) and the
conditional distribution of \(\Theta\) given \(Y=y\) is a sine skewed
von Mises distribution with location parameter \(\mu\) and concentration
parameter \((\nu y)^\alpha\tanh(\kappa)\). The log-likelihood for this
model equals \begin{align}\label{WeiSSVMLikelihood}
l(\alpha, \boldsymbol{\gamma}, \lambda, \kappa, \boldsymbol{\beta}) 
   &= n[\ln(\alpha) - \ln(2\pi\cosh(\kappa))] + \alpha \sum^{n}_{i = 1} \boldsymbol{x}_i^t\boldsymbol{\gamma} \nonumber\\
   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - (\beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})))) 
   +(\alpha-1)\sum^{n}_{i = 1} \ln(y_i) \nonumber\\
   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - (\beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})))).\nonumber
\end{align} \noindent We can use numerical optimization (Nelder-Mead) to
find solutions for the maximum likelihood (ML) estimates for the
parameters of the model.

@@

intercept and the \(g\) covariate values. \newline

@@

\end{enumerate}
\newpage
\section*{References}
